---
title: "Creating a file of a pre-mapped nameset for OpenTree's import nameset tool"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating a file of a pre-mapped nameset for OpenTree's import nameset tool}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library(ropentree)
```

## Creating a CSV file of a pre-mapped nameset

If your taxon names are on a newick tree file, read the tree into R with the `read.tree` function from the `ape` package. This will create a `phylo` object:

```{r}
phy <- ape::read.tree(file="../data-raw/mgharvey-tyranni-f73aa7f/species_trees/final_timetrees/T400F_complete.tre")
ls(phy)
str(phy)
```

Now, create a vector of original tip taxon labels

```{r}
original_labels <- phy$tip.label

head(original_labels, 10)
```

In this particular case, tip taxon labels are not suitable for mapping because the scientific names are cut off (e.g. Homo_sapiens is Hom_sap_000).
So, we will need to link the tip taxon labels with the complete scientific name, as provided on a supplementary data table from the original publication.

To do this, first read the supplementary data table file linking tip taxon labels with the complete scientific taxon names using the `read.csv` function:

```{r}
data_table <- read.csv(file = "../data-raw/mgharvey-tyranni-f73aa7f/Species_name_map_uids.csv")

head(data_table, 10)
ls(data_table)
class(data_table)
```

Now, we can create a vector of scientific species names that are suitable for TNRS matching.
To save on computing time, let's use the `unique` function to eliminate duplicated scientific names and leave just one copy per name:

```{r}
length(data_table$matrix.species)
adjusted_labels <- unique(data_table$matrix.species)
head(adjusted_labels, 10)
length(adjusted_labels)
```
Let's check for NA's and remove them:

```{r}
any(is.na(adjusted_labels))
index_na <- which(is.na(adjusted_labels))
adjusted_labels <- adjusted_labels[-index_na]
```

For the sake of reproducibility, write down the names as a ".txt" file that can be read back into R using `read.csv`:

```{r, eval = FALSE}

write(adjusted_labels, file = "../data-raw/harvey2020scientific_names.txt")

adjusted_labels  <-  read.csv(file = "../data-raw/harvey2020scientific_names.txt", header = FALSE)[,1]
```

Now, perform a TNRS match on the vector of unique scientific names with the `tnrs_match_names` function from the `rotl` package:

```{r}
matched_labels  <-  sapply(adjusted_labels, function(x){
  write(x, file = "../data-raw/test.txt", append = TRUE)
  rotl::tnrs_match_names(names = x)}
  )
head(matched_labels, 10)
```

Now, we need to extract the OTT ids that resulted from the TNRS match.
First, match the original vector of tip taxon names with results from the TNRS match:

```{r}
index <- match(data_table$matrix.species, names(matched_labels["ott_id",]))
head(index, 10)
```

This will allow you to create a vector of OTT ids that correspond to the original tip taxon labels: 

```{r}
ott <- unlist(matched_labels["ott_id",][index])
head(ott, 10)
```

Finally, create a data frame containing the necessary information for OpenTree's importing nameset tool, and write it as csv file:

```{r, eval=FALSE}
results <- data.frame(original = original_labels, adjusted = data_table$matrix.species, ott_id = ott)

write.csv(results, file = "../data-raw/harvey2020nameset.csv)
```


## Creating a json file 


