---
title: "plant_genera"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rotl)
```

# Get a set of taxon names

Read a file with taxon names. The function `read.csv` will read them as a data frame or table.

```{r}
taxa <- read.csv("../data-raw/plant_genera.txt", header = FALSE)
str(taxa)
```

Make a character vector of taxon names
```{r}

taxa <- taxa$V1
str(taxa)
```

# Match taxon names to OpenTree Taxonomy (OTT)

We will use the `tnrs_match_names` function from the `rotl` R package.
```{r}
taxa_tnrs <- rotl::tnrs_match_names(taxa)
ls(taxa_tnrs)
```
It generates a dat frame with teh results of the TNRS match to the OpenTree Taxonomy (OTT).

We want the OTT id numbers.

```{r}
taxa_ott_ids <- taxa_tnrs$ott_id
```

# Get an OpenTree synthetic subtree for each genus

To extract a synthetic subtree containing all descendants of any taxon, we will use the function `tol_subtree`, from the `rotl` R package, again.

```{r}
subtree_taxon1 <- rotl::tol_subtree(taxa_ott_ids[1], label_format = "name")
subtree_taxon1
```

The function just allows one OTT id at a time, to get all subtrees at once, we can use a for loop or a handy `lapply` function.

```{r}
subtree_taxa_all_name <- lapply(taxa_ott_ids, rotl::tol_subtree, label_format = "")
subtree_taxa_all_id <- lapply(taxa_ott_ids, rotl::tol_subtree, label_format = "id")
```
It is easier to catch errors with a for loop:

```{r}
subtree_taxa_all_id <- vector(mode = "list")
for (n in taxa_ott_ids){
  subtree <- try(rotl::tol_subtree(n, label_format = "id"))
  # print(subtree)
  subtree_taxa_all_id <- c(subtree_taxa_all_id, list(subtree))
  
}
length(subtree_taxa_all_id)
```

Let's do the same with scientific names:

```{r}
subtree_taxa_all_name <- vector(mode = "list")
for (n in taxa_ott_ids){
  subtree <- try(rotl::tol_subtree(n, label_format = "name"))
  # print(subtree)
  subtree_taxa_all_name <- c(subtree_taxa_all_name, list(subtree))
  
}
length(subtree_taxa_all_name)
```
Check which ones errored and which ones produced a subtree:

```{r}
sapply(subtree_taxa_all, class)

is_error <-   sapply(subtree_taxa_all, class) %in% "try-error"
is_phylo <-   sapply(subtree_taxa_all, class) %in% "phylo"
```

Out of the `r length(subtree_taxa_all)` taxon OTT ids tried, `r sum(is_error)` were unsuccessful and the remainder `r sum(is_phylo)` retrieved a subtree successfully.

We will make a new list object containing subtrees only, excluding the errors:

```{r}
subtree_taxa_phylo_name <- subtree_taxa_all_name[is_phylo]
length(subtree_taxa_phylo_name)
names(subtree_taxa_phylo_name) <- taxa[is_phylo]

subtree_taxa_phylo_id <- subtree_taxa_all_id[is_phylo]
length(subtree_taxa_phylo_id)
names(subtree_taxa_phylo_id) <- taxa[is_phylo]

```
Now, we can make a vector containing all taxon names that are tip labels in the synthetic subtrees:

```{r}
subtaxa_name <- sapply(subtree_taxa_phylo_name, "[", "tip.label")
length(subtaxa_name)
names(subtaxa_name) <- names(subtree_taxa_phylo_name)

subtaxa_id <- sapply(subtree_taxa_phylo_id, "[", "tip.label")
length(subtaxa_id)
names(subtaxa_id) <- names(subtree_taxa_phylo_id)
subtaxa_id[1]
```


# Get GBIF taxon names of tree tip labels (necessary??)

# Get GBIF data for tips in tree

We will use the function `occ_search` from the `rgbif` R package.
Let's try it first with the scientific names in `subtaxa_name`:

```{r}

name <- gsub("_", " ", subtaxa_name[[1]][1])
name_gbif <- rgbif::occ_search(scientificName = name, fields=c('name','decimalLatitude', 'decimalLongitude'), limit = 100)
ls(name_gbif)
ls(name_gbif$data)
name_gbif$data
```

Next, we will do it in a loop:

```{r}
# for each phylo object i
for (i in 1:length(subtaxa_name)){
  print(i)
  # create an empty list to hold results
  subtaxa_gbif <- vector(mode = "list")
  # for each tip label "name" in phylo object i
  for (name in subtaxa_name[[i]]){
    print(name)
    # replace underscores by spaces in name
    name <- gsub("_", " ", name)
    # get gbif records for "name", with a try
    name_gbif <- try(rgbif::occ_search(scientificName = name, 
                                   fields=c('order',
                                            'family',
                                            'genus',
                                            'acceptedScientificName',
                                            'decimalLatitude', 
                                            'decimalLongitude', 
                                            "elevation", 
                                            "familyKey", 
                                            "genusKey", 
                                            "taxonKey"),
                                   limit = 5))
    # if try was successful and produced a gbif object
    # and data is not empty (i.e., there are gbif records for "name")
    if(inherits(name_gbif, "gbif") & !is.null(name_gbif$data)){
      # sometimes, there are records but they are not retrieved
      # if data was more than 0 rows
      if(nrow(name_gbif$data)){
        # create a data frame object containing "name" and gbif data for "name"
        name_gbif <- cbind(name, name_gbif$data)
        
      }
    }
  # concatenate the result of each "name" into a list
  subtaxa_gbif <- c(subtaxa_gbif, list(name_gbif))
  }
  # use subtaxon names to name the list of results for all subtaxa in taxa
  names(subtaxa_gbif) <- subtaxa_name[[i]]
  # save results as an R object with name = names(subtaxa_name)[i]
  assign(names(subtaxa_name)[i], get("subtaxa_gbif"))
  save(list = names(subtaxa_name)[i], file = paste0("../data-raw/", names(subtaxa_name)[i], ".RData"))
  # which ones are data frames
  is_data <- sapply(subtaxa_gbif, class) %in% "data.frame"
  # merge all data into a single table
  tab <- dplyr::bind_rows(subtaxa_gbif[is_data])
  write.csv(tab, file = paste0("../data-raw/", names(subtaxa_name)[i], ".csv"))
}

```

```{r}
